Mining Scenario/Transaction Pairs in HW/SW Co-verification Using Symbolic Execution

### 2017-04-07 paper.tex
fixes
minor

\subsection{Approach 1: Path-based Co-verification Technique}

@@

\textbf{Step 4: Property Driven Slicing of Co-verification Model}
A property-driven slicing is performed on the unified model, $\mathcal{M}_{seq}$.  
This step is purely syntactic, meaning that we perform a backward dependency 
analysis starting from the property which only preserve those 
program fragments that are relevant to a given property. The sliced program 
is then passed to the symbolic execution engine for path-based exploration. \\
%
%=======================================================================================
%\subsection{\verifox: Path-wise Symbolic Execution Tool} 
%=======================================================================================
\input{verifox}
%
\subsection{Approach 2: Monolithic Co-verification Technique}
%
\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.6]{figures/traditional_flow.pdf}
\caption{\small Monolithic HW/SW Co-verification Flow in \hwcbmcv}
\label{fig:conventional}
\end{center}
\end{figure}
%
\textbf{Step 1: Co-verification Model}
%
Figure~\ref{fig:conventional} illustrates the monolithic approach to HW/SW
co-veri\-fi\-cation, as implemented in \hwcbmcv.  
%\hwcbmcv is the only open-source
%formal co-verification tool that supports Verilog RTL (System Verilog 2009
%standard) and C/C++ firmware.  
In contrast to path-based approach,
\hwcbmcv maintains two separate flows for HW and FW.  The top flow in
Figure~\ref{fig:conventional} uses synthesis to obtain either a bit-level or
a word-level netlist from Verilog RTL.  A bit-level netlist is represented 
in And Inverter Graph (AIG) format.  Whereas, a word-level netlist is represented 
in format that resembles SMT-LIB2 standards~\footnote{http://smtlib.cs.uiowa.edu/}.  
The bottom flow illustrates the translation of C program 
into static single assignment (SSA) form. 
Synthesizing netlist from RTL is a standard technique which is widely 
used in hardware model checkers~\cite{DBLP:journals/tse/Clarke76}.
Similarly, translating a C program into an SSA representation is a 
standard technique which is widely used in 
software verification tools, such as CBMC~\cite{cbmc.tacas:2004}. 
\hwcbmcv generates a monolithic co-verification model from AIG and SSA through 
in-tandem symbolic execution of HW and FW components.

@@

\textbf{Step 3: Scenario/Transaction Pair} In \hwcbmcv, the symbolic execution 
of HW and FW models are clearly separate and the two flow meet only at the
solver phase, where a complex monolithic formula is generated in 
Conjunctive Normal Form (CNF) from the AIG and SSA.  This complex formula 
is passed on to the solver for verification purpose.  
Hence, the verification flow in \hwcbmcv limits the opportunity for extracting 
relevant scenario/transaction pairs through symbolic execution.  Instead, the 
technique relies on the underlying SAT/SMT solver to efficiently reason about 
the co-verification model. 
%
\Omit{
The hardware and firmware run independently of each other. 
The communication between them takes place through these function calls. 
}
%===============================================================================
\section{Property Specification for Co-verification}\label{property}
%===============================================================================
Lack of support for property specification in a HW/SW co-design is one of the
stumbling blocks for the application of formal techniques in co-verification.  
In this paper, we express a co-design property in C 
with the support of special primitive, called {\em next\_timeframe()}, 
that can express temporal correlation between HW and FW events. 
In this paper, we classify co-design properties into two 
classes -- {\em transaction level} property and {\em component level} property.
%-------------------------------------------------------------------------------
\subsection{Transaction level Property}
%-------------------------------------------------------------------------------
A {\em Transaction level} property ($\psi_{t}$) formally captures the 
design intent of the interaction between a $\mathcal{FT}$ and $\mathcal{HT}$, 
that is, it is defined at the transaction boundaries.  
Recall that a $\mathcal{FT}$ is a service in response to some external environment 
or HW interrupts (for example acknowledgement or exceptions). Similarly, 
$\mathcal{HT}$ is a finite-length service in response to some SW events or 
other HW events.

@@

%-------------------------------------------------------------------------------
\subsection{Component level Property}
%-------------------------------------------------------------------------------
A {\em component level} property ($\psi_{c}$) is defined purely at the 
level of individual components of a co-design system. Recall that a component 
may be a HW model, a SW model or an interface model which contains only the 
interface logic.  A component-level property captures the design intent of a 
component in terms of its input/output behaviors.  This type of property 
is useful in a co-design environment for verifying individual components.   
Note that unlike the structure of $\psi_{t}$ which requires the antecedent 
or consequent to be a HW or SW event,  $\psi_{c}$ only contains signals 
(ports) or variables of a particular component.  Figure~\ref{figure:component} 
shows an example of component level property for the UART HW only.  The 
System Verilog Assertion (SVA) is shown on the left and the equivalent 
monitor in C is shown on the right.  
%
\begin{figure}[htbp]
\scriptsize
\begin{tabular}{l|l}
\hline
System Verilog Assertions  & Monitor (in C)
\\
\hline
\begin{lstlisting}[mathescape=true,language=Verilog]
P3: assert property 
 (@(posedge clk) 
 ack == 1|-> 
 (valid == 1 && 
 ##2 empty == 0)); 
\end{lstlisting}
&
\begin{lstlisting}[mathescape=true,language=C]
bool monitor_P3() {
 assert(!ack || valid);
 next_timeframe();
 next_timeframe();
 assert(empty==0); }
\end{lstlisting} \\
\hline
\end{tabular}
\caption{Component level property as monitor (in C)}
\label{figure:component}
\end{figure}
%
%-------------------------------------------------------------------------------
\subsection{Debugging Using Counterexample Trace}
%-------------------------------------------------------------------------------
A failure of co-design property implies a fault in HW or a defective FW 
or bug in the interface functions.  Hence, an important feature of a
co-verification tool is its ability to produce a counterexample trace 
in order to localize the bug.  \hwcbmcv generates a counterexample
trace in value change dump (vcd) format that contains concrete assignments to FW
variables, interface signals and HW RTL signals up to a given bound. Whereas, a
counterexample trace in \verifox is a bounded path in the sequential co-verification
model that contains a scenario and its corresponding HW and FW transactions. 
%
%===============================================================================
\section{Experimental Results}
%===============================================================================
%
We report experimental results for FW-HW co-verification of a UART  and a
SoC design.  All our experiments were performed on an Intel Xeon 3.07\,GHz
machine with 48\,GB RAM.  All times reported are in seconds.  The timeout 
for all our experiments was set to 2~hours.  
%

@@

the verification outcome. The focus of our experiment is to compare 
the performance of path-based technique with monolithic BMC approach for 
co-verification.  We observe similar performance of bit-level and word-level 
flow in \hwcbmcv. So, we only report bit-level results for \hwcbmcv in this
paper. \\ 
%\sjcmt{I think this can backfire. The reivewers may ask why were other results not reported if available}\\ 

@@

\begin{table}
\begin{center}
{
\begin{scriptsize}
\begin{tabular}{|l|l|l|l|l|l|l|}
\hline
  & & \multicolumn{1}{c|}{Monolithic} & \multicolumn{3}{c|}{Path-based} &
  Verification \\ 
\cline{3-6}
  Scenario & Bound & \multicolumn{1}{c|}{HW-CBMC} & \multicolumn{3}{c|}{\verifox} &
  Results \\ 
\cline{3-7}
      &       &  Bit-level & Total/Feasible & \%-age & $\mathcal{M}_{seq}$ &
      Safe/Unsafe \\
  \cline{3-3}\cline{6-6}
      &       &   Time     & Paths & Pruning & Time &  \\
\cline{1-7}      
\multicolumn{7}{|c|}{non-deterministic data but deterministic control (Scenario A)} \\ \hline
  transmit (32) & 250 & 15.02 & 247104/224 & 99.90 & \textbf{1}.\textbf{13} &
  Safe $(\psi_{c})$ \\ 
transmit (64) & 500 & 23.87 & 247104/324 & 99.86 &
  \textbf{1}.\textbf{61}  & Safe $(\psi_{t})$ \\ \hline 
\multicolumn{7}{|c|}{non-deterministic data and non-deterministic control
(Scenario B)} \\ \hline
  trans\_intr (32) & 250 & 14.86 & 247104/295 & 99.88 & \textbf{1}.\textbf{49} &
  Safe $(\psi_{t})$ \\
  trans\_intr (64) & 500 & 24.14 & 247104/362 & 99.85 & \textbf{1}.\textbf{81}
  & Safe $(\psi_{t})$ \\ \hline
\multicolumn{7}{|c|}{non-deterministic data and non-deterministic control
(scenario C)} \\ \hline
  loopback (8)  & 230 & 52.06 & 247104/354 & 99.85 & \textbf{3}.\textbf{95} &
  Safe $(\psi_{t})$ \\ 
  loopback (16) & 500 & 122.12 & 247104/690 & 99.72 & \textbf{12}.\textbf{89} &
  Safe $(\psi_{c})$ \\ 
  loopback (32) & 650 & 170.62 & 247104/1282 & 99.48 & \textbf{21}.\textbf{85} &
  Safe $(\psi_{c})$ \\ 
  loopback (64) & 1300 & 409.71 & 247104/2566 & 98.96 & \textbf{62}.\textbf{31}
  & Safe $(\psi_{t})$ \\ \hline
\multicolumn{7}{|c|}{detecting data-path bugs in transmission mode w/o interrupt} \\ \hline
  transmit (64) & 520 & 28.43 & 247104/324 & 99.86 &
  \textbf{1}.\textbf{12} & Unsafe $(\psi_{t})$ \\ \hline
\multicolumn{7}{|c|}{detecting control bugs with interrupt enabled} \\ \hline
transmit (64) & 520 & 31.35 & 247104/362 & 99.85 &
  \textbf{1}.\textbf{05} & Unsafe $(\psi_{c})$ \\ \hline
\multicolumn{7}{|c|}{detecting control bugs in loopback mode} \\ \hline
  loopback (64) & 1300 & 443.15 & 247104/2566 & 98.96 & \textbf{62}.\textbf{34}
  & Unsafe $(\psi_{t})$ \\ \hline
\end{tabular}
\end{scriptsize}
}
\end{center}
%\vspace{-1.3mm}
\caption{Verification times for UART design
\label{table:safe}}
\end{table}

@@

      &       &   Time      & Time &  \\

@@

\caption{Verification times for SoC Design

@@

\Omit{
The concept of symbolic execution~\cite{DBLP:journals/tse/Clarke76,
DBLP:conf/pldi/GodefroidKS05, DBLP:conf/osdi/CadarDE08} is prevalent in the
software domain for automated test generation as well as bug finding.  This
technique is different from the symbolic simulation techniques that are used
in the hardware domain.  In this paper, we apply path-wise symbolic execution
to mine relevant scenario/transaction pairs for effective co-verification of 
a unified co-verification model.  
}
%===============================================================================
\section{Concluding Remarks}
%===============================================================================
%
In this paper, we presented two different approaches for formal verification of 
HW/SW co-designs -- \emph{monolithic} and \emph{path-based}.  These approaches are 
embodied in our tools, \hwcbmcv and \verifox, respectively.  Our tools support 
co-designs written in Verilog RTL and C.  We also presented a specification language to 
effectively capture co-design properties.  
%To this end, we classify co-design properties into two types --  \emph{transaction level} and \emph{component level} property.  
We proved several transaction and component level properties of UART and SoC
design using \hwcbmcv and \verifox.  
%We for proving safety as well as finding critical co-design bugs for a UART IP and a complex SoC design.  
We uncovered three critical bugs in UART and one control-path bug in the SoC 
within only couple of minutes.

In practice, a FW exercises only a fragment of the HW state-space for a given scenario.  
This renders many interactions between HW and FW transactions infeasible.   
Our general observation is that bounded model checking technique in \hwcbmcv 
cannot prune irrelevant transactions with respect to a given scenario, 
and hence generate a monolithic co-verification model, which are often 
extremely difficult to solve with a SAT/SMT solver. On the other hand, 
a path-based exploration strategy in \verifox is able to automatically 
mine relevant scenario/transaction pairs in a sequential co-verification 
model, owing to specific optimizations such as eager path pruning
combined with incremental SAT solving.  Our experiments show that \verifox 
is on average $5\times$ faster than \hwcbmcv for proving safety as well as 
bug finding.  In the future, we plan to extend our framework to co-design 
models that exhibit further interaction patterns.

>>>>

\subsection{Approach 1: Path-based Co-verification}

@@

\textbf{Step 4: Property Driven Slicing of Co-verification Model}
A property-driven slicing is performed on the unified model, $\mathcal{M}_{seq}$.  
This step is purely syntactic, meaning that we perform a backward dependency 
analysis starting from the property which only preserve those 
program fragments that are relevant to a given property. The sliced program 
is then passed to the symbolic execution engine for path-based exploration. \\
%
%=======================================================================================
%\subsection{\verifox: Path-wise Symbolic Execution Tool} 
%=======================================================================================
\input{verifox}
%
\subsection{Approach 2: Monolithic Co-verification}
%
\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.6]{figures/traditional_flow.pdf}
\caption{\small Monolithic HW/SW Co-verification Flow in \hwcbmcv}
\label{fig:conventional}
\end{center}
\end{figure}
%
\textbf{Step 1: Co-verification Model}
%
Figure~\ref{fig:conventional} illustrates the monolithic approach to HW/SW
co-veri\-fi\-cation, as implemented in \hwcbmcv.  
%\hwcbmcv is the only open-source
%formal co-verification tool that supports Verilog RTL (System Verilog 2009
%standard) and C/C++ firmware.  
In contrast to path-based approach,
\hwcbmcv maintains two separate flows for HW and FW.  The top flow in
Figure~\ref{fig:conventional} uses synthesis to obtain either a bit-level or
a word-level netlist from Verilog RTL.  A bit-level netlist is represented 
in And Inverter Graph (AIG) format.  Whereas, a word-level netlist is represented 
in format that resembles SMT-LIB2 standards~\footnote{http://smtlib.cs.uiowa.edu/}.  
The bottom flow illustrates the translation of C program 
into static single assignment (SSA) form. 
Synthesizing netlist from RTL is a standard technique which is widely 
used in hardware model checkers~\cite{DBLP:journals/tse/Clarke76}.
Similarly, translating a C program into an SSA representation is a 
standard technique which is widely used in 
software verification tools, such as CBMC~\cite{cbmc.tacas:2004}. 
\hwcbmcv generates a monolithic co-verification model from AIG and SSA through 
in-tandem symbolic execution of HW and FW components.

@@

\textbf{Step 3: Scenario/Transaction Pair} In \hwcbmcv, the symbolic execution 
of HW and FW models are clearly separate and the two flow meet only at the
solver phase, where a complex monolithic formula is generated in 
Conjunctive Normal Form (CNF) from the AIG and SSA.  This complex formula 
is passed on to the solver for verification purpose.  
Hence, the verification flow in \hwcbmcv limits the opportunity for extracting 
relevant scenario/transaction pairs through symbolic execution.  Instead, the 
technique relies on the underlying SAT/SMT solver to efficiently reason about 
the co-verification model. 
%
\Omit{
The hardware and firmware run independently of each other. 
The communication between them takes place through these function calls. 
}
%===============================================================================
\section{Co-verification Property}\label{property}
%===============================================================================
Lack of support for property specification in a HW/SW co-design is one of the
stumbling blocks for the application of formal techniques in co-verification.  
In this paper, we express a co-design property in C 
with the support of special primitive, called {\em next\_timeframe()}, 
that can express temporal correlation between HW and FW events. 
In this paper, we classify co-design properties into two 
classes -- {\em transaction level} property and {\em component level} property.
%-------------------------------------------------------------------------------
\subsection{Transaction level Property}
%-------------------------------------------------------------------------------
A {\em Transaction level} property ($\psi_{t}$) formally captures the 
design intent of the interaction between a $\mathcal{FT}$ and $\mathcal{HT}$, 
that is, it is defined at the transaction boundaries.  
Recall that a $\mathcal{FT}$ is a service in response to some external environment 
or HW interrupts (for example acknowledgement or exceptions). Similarly, 
$\mathcal{HT}$ is a finite-length service in response to some SW events or 
other HW events.

@@

%-------------------------------------------------------------------------------
\subsection{Component level Property}
%-------------------------------------------------------------------------------
A {\em component level} property ($\psi_{c}$) is defined purely at the 
level of individual components of a co-design system. Recall that a component 
may be a HW model, a SW model or an interface model which contains only the 
interface logic.  A component-level property captures the design intent of a 
component in terms of its input/output behaviors.  This type of property 
is useful in a co-design environment for verifying individual components.   
Note that unlike the structure of $\psi_{t}$ which requires the antecedent 
or consequent to be a HW or SW event,  $\psi_{c}$ only contains signals 
(ports) or variables of a particular component.  Figure~\ref{figure:component} 
shows an example of component level property for the UART HW only.  The 
System Verilog Assertion (SVA) is shown on the left and the equivalent 
monitor in C is shown on the right.  
%
\begin{figure}[htbp]
\scriptsize
\begin{tabular}{l|l}
\hline
System Verilog Assertions  & Monitor (in C)
\\
\hline
\begin{lstlisting}[mathescape=true,language=Verilog]
P3: assert property 
 (@(posedge clk) 
 ack == 1|-> 
 (valid == 1 && 
 ##2 empty == 0)); 
\end{lstlisting}
&
\begin{lstlisting}[mathescape=true,language=C]
bool monitor_P3() {
 assert(!ack || valid);
 next_timeframe();
 next_timeframe();
 assert(empty==0); }
\end{lstlisting} \\
\hline
\end{tabular}
\caption{Component level property as monitor (in C)}
\label{figure:component}
\end{figure}
%
%-------------------------------------------------------------------------------
\subsection{Debugging Using Counterexample Trace}
%-------------------------------------------------------------------------------
A failure of co-design property implies a fault in HW or a defective FW 
or bug in the interface functions.  Hence, an important feature of a
co-verification tool is its ability to produce a counterexample trace 
in order to localize the bug.  \hwcbmcv generates a counterexample
trace in value change dump (vcd) format that contains concrete assignments to FW
variables, interface signals and HW RTL signals up to a given bound. Whereas, a
counterexample trace in \verifox is a bounded path in the sequential co-verification
model that contains a scenario and its corresponding HW and FW transactions. 
%
%===============================================================================
\section{Experimental Results}
%===============================================================================
%
We report experimental results for FW-HW co-verification of a UART  and a
SoC design.  All our experiments were performed on an Intel Xeon 3.07\,GHz
machine with 48\,GB RAM.  All times reported are in seconds.  The timeout 
for all our experiments was set to 2~hours.  The performance of bit-level 
and word-level flow in \hwcbmcv are similar. So, we only report bit-level 
results for \hwcbmcv.  MiniSAT-2.2.0~\cite{DBLP:conf/sat/EenB05} was used 
as underlying SAT solver with \hwcbmcv and \verifox.  The focus of our 
experiments is to compare the performance of path-based technique in \verifox 
with monolithic BMC approach in \hwcbmcv for verification of UART and SoC design.  
\\ 
%

@@

the verification outcome.\\ 

@@

\begin{table}
\begin{center}
{
\begin{scriptsize}
\begin{tabular}{|l|l|l|l|l|l|l|}
\hline
  & & \multicolumn{1}{c|}{Monolithic} & \multicolumn{3}{c|}{Path-based} &
  Verification \\ 
\cline{3-6}
  Scenario & Bound & \multicolumn{1}{c|}{HW-CBMC} & \multicolumn{3}{c|}{\verifox} &
  Results \\ 
\cline{3-7}
      &       &  Bit-level & Total/Feasible & \%-age & $\mathcal{M}_{seq}$ &
      Safe/Unsafe \\
  \cline{3-3}\cline{6-6}
      &       &   Time     & Paths & Pruning & Time &  \\
\cline{1-7}      
\multicolumn{7}{|c|}{non-deterministic data but deterministic control (Scenario A)} \\ \hline
  transmit (32) & 250 & 15.02 & 247104/224 & 99.90 & \textbf{1}.\textbf{13} &
  Safe $(\psi_{c})$ \\ 
transmit (64) & 500 & 23.87 & 247104/324 & 99.86 &
  \textbf{1}.\textbf{61}  & Safe $(\psi_{t})$ \\ \hline 
\multicolumn{7}{|c|}{non-deterministic data and non-deterministic control
(Scenario B)} \\ \hline
  trans\_intr (32) & 250 & 14.86 & 247104/295 & 99.88 & \textbf{1}.\textbf{49} &
  Safe $(\psi_{t})$ \\
  trans\_intr (64) & 500 & 24.14 & 247104/362 & 99.85 & \textbf{1}.\textbf{81}
  & Safe $(\psi_{t})$ \\ \hline
\multicolumn{7}{|c|}{non-deterministic data and non-deterministic control
(scenario C)} \\ \hline
  loopback (8)  & 230 & 52.06 & 247104/354 & 99.85 & \textbf{3}.\textbf{95} &
  Safe $(\psi_{t})$ \\ 
  loopback (16) & 500 & 122.12 & 247104/690 & 99.72 & \textbf{12}.\textbf{89} &
  Safe $(\psi_{c})$ \\ 
  loopback (32) & 650 & 170.62 & 247104/1282 & 99.48 & \textbf{21}.\textbf{85} &
  Safe $(\psi_{c})$ \\ 
  loopback (64) & 1300 & 409.71 & 247104/2566 & 98.96 & \textbf{62}.\textbf{31}
  & Safe $(\psi_{t})$ \\ \hline
\multicolumn{7}{|c|}{detecting data-path bugs in transmission mode w/o interrupt} \\ \hline
  transmit (64) & 520 & 28.43 & 247104/324 & 99.86 &
  \textbf{1}.\textbf{12} & Unsafe $(\psi_{t})$ \\ \hline
\multicolumn{7}{|c|}{detecting control bugs with interrupt enabled} \\ \hline
transmit (64) & 520 & 31.35 & 247104/362 & 99.85 &
  \textbf{1}.\textbf{05} & Unsafe $(\psi_{c})$ \\ \hline
\multicolumn{7}{|c|}{detecting control bugs in loopback mode} \\ \hline
  loopback (64) & 1300 & 443.15 & 247104/2566 & 98.96 & \textbf{62}.\textbf{34}
  & Unsafe $(\psi_{t})$ \\ \hline
\end{tabular}
\end{scriptsize}
}
\end{center}
%\vspace{-1.3mm}
  \caption{Verification of UART (All time in seconds)
\label{table:safe}}
\end{table}

@@

      &       &   Time (seconds)       & Time (seconds) &  \\

@@

  \caption{Verification times for SoC Design (All time in Seconds)

@@

\Omit{
The concept of symbolic execution~\cite{DBLP:journals/tse/Clarke76,
DBLP:conf/pldi/GodefroidKS05, DBLP:conf/osdi/CadarDE08} is prevalent in the
software domain for automated test generation as well as bug finding.  This
technique is different from the symbolic simulation techniques that are used
in the hardware domain.  In this paper, we apply path-wise symbolic execution
to mine relevant scenario/transaction pairs for effective co-verification of 
a unified co-verification model.  
}
%===============================================================================
\section{Concluding Remarks}
%===============================================================================
%
In this paper, we presented two different approaches for formal verification of 
HW/SW co-designs -- \emph{monolithic} and \emph{path-based}.  These approaches are 
embodied in our tools, \hwcbmcv and \verifox, respectively.  Our tools support 
co-designs written in Verilog RTL and C.  We also presented a specification language to 
effectively capture co-design properties.  We proved several transaction and component 
level properties of UART and SoC design using \hwcbmcv and \verifox.  We uncovered three 
critical bugs in UART and one control-path bug in the SoC within only couple of minutes.

In practice, a FW exercises only a fragment of the HW state-space for a given scenario.  
This renders many interactions between HW and FW transactions infeasible.   
Our general observation is that bounded model checking technique in \hwcbmcv 
cannot prune irrelevant transactions with respect to a given scenario, 
and hence generate a monolithic co-verification model, which are often 
extremely difficult to solve with a SAT/SMT solver. On the other hand, 
a path-based exploration strategy in \verifox is able to automatically 
mine relevant scenario/transaction pairs in a sequential co-verification 
model, owing to specific optimizations such as eager path pruning
combined with incremental SAT solving.  Our experiments show that \verifox 
is on average $5\times$ faster than \hwcbmcv for proving safety as well as bug
finding.  In the future, we plan to extend our framework to co-design 
that exhibit further interaction patterns.
